---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
>Note to self:
Markdown doesn't show the Python code in preview, only shows the output of the Python code.
Might be more sensible to just use a normal R script for the demo



# Reticulate: Unifying R and Python

[Stuff here about not having to choose between R and Python]

## 0. Environment set-up

First, let's load reticulate and set up our Conda environment. Running ```use_condaenv``` will activate the specified environment, which may first need to be created either using ```conda_create``` or your Conda management tool of choice. Note that Reticulate installs its own Miniconda instance, so beware of this if managing Conda environments from outside of R!

To learn more about Conda environments, please see [this link](https://docs.conda.io/projects/conda/en/latest/user-guide/concepts/environments.html).


```{r}
# R
library(reticulate)

env_name <- "sklearn_demo"
env_list <- conda_list()$name
if (!env_name %in% env_list){
  conda_create(envname=env_name, packages=c('python', 'scikit-learn', 'pandas'))
}

use_condaenv(env_name, required=TRUE)
```

## 1. Setting and accessing variables

```{r}
# R
greeting <- "Hello world, from R"
```

R variables are accessed in Python through the ```r``` object's attributes.

```{python}
# Python
greeting = "Hello world, from Python"

print(r.greeting)
```

Similarly, we access our Python variables in R with the ```py``` object. 

```{r}
# R
print(py$greeting)
```

We can set Python variables from within R:

```{r}
# R
py$new_var <- 'This variable was set in R'
```
```{python}
# Python
print(new_var)
```

Or vice-versa:

```{python}
# Python
r.new_var = 'This variable was set in Python'
```
```{r}
# R
print(new_var)
```

## 2. Data handling

```{r}
# R
iris <- datasets::iris
head(iris)
```
```{python}
# Python

import pandas
```

```{r}
# R
library(ggplot2)

ggplot(data=iris, aes(x=Petal.Length, y=Petal.Width, color=Species)) + 
  geom_point(aes(shape=Species), size=1.5) + xlab("Petal Length") + ylab("Petal Width") +   
  geom_smooth(method="lm")
```


```{python}
# Python
from sklearn import svm

r.iris
```


